---
title: "Hypothesis Testing"
author: "Marie Khalil"
date: "`r Sys.Date()`"
output: html_document
---
Loading libraries

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(janitor)
```

Loading and Viewing the first 6 rows of the table

```{r}
activity_sleep_weight <- read_csv("./Data/cleaned_data/activity_sleep_weight_daily_joined_08_05_2022_v02.csv")
head(activity_sleep_weight)
```

number of active users per day


```{r}
users_per_day <- activity_sleep_weight %>%
  group_by(activity_date) %>% 
  summarise(users_count = n())

users_per_day

```

Checking the distribution of data 

```{r}

mean_users_count = mean(users_per_day$users_count)
sd_users_count = sd(users_per_day$users_count)

x__ = rnorm(31, mean = mean_users_count, sd = sd_users_count)

ggplot(data = users_per_day) + 
  geom_histogram(aes(x= as.integer(users_count)), binwidth = .5)+
  geom_density(aes(x = x__ ), colour = "red", show.legend = FALSE)+ 
  geom_point(aes(x = mean_users_count, y = 0 ), colour = "red")+
  geom_point(aes(x = mean_users_count + sd_users_count, y = 0), colour = "green")+
  geom_point(aes(x = mean_users_count - sd_users_count, y = 0), colour = "green")+
  xlim(9,18)+
  xlab(" Number of active users")+
  ylab("Frequency")+
  ggtitle("Distribution of Active users per day",
          subtitle = "N = 24 \n Duration: 13 Apr to 13 May")
```

```{r}
ggplot(data = users_per_day) +
  geom_density(aes(x = as.integer(users_count) ), colour = "red", show.legend = FALSE)+
  xlim(9,18)+
  xlab(" Number of active users")+
  ylab("Density")+
  ggtitle("Density of Active users per day",
          subtitle = "N = 24 \n Duration: 13 Apr to 13 May")
```

according to the density plot, the number of active users per day is
normally distributed (bell shaped curve)

we can proceed to check if there is correlation between **weekdays** and
**number of Active users**

adding new column weekday and converting the weekday column to be a
factor with predefined levels then order the table be weekdays starting
from Monday

```{r}
activity_sleep_weight$weekday <- weekdays(activity_sleep_weight$activity_date)
activity_sleep_weight$weekday <- factor(activity_sleep_weight$weekday, levels = c("Monday", 
    "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday","Sunday"))

activity_sleep_weight_orderd <- activity_sleep_weight[order(activity_sleep_weight$weekday),]
activity_sleep_weight_orderd
```

```{r}
weekdays_df <- activity_sleep_weight_orderd %>%
  group_by(activity_date,weekday) %>% summarise(user_count= n())
weekdays_df 
```

boxplot

```{r}
ggplot(data=weekdays_df)+ 
  geom_boxplot(aes(x=weekday,y = user_count),
               outlier.colour="red",
               outlier.shape=8,
               outlier.size=4)+
  ylim(7,18)
```

according to the boxplot 

* Tuesday and Thursday have the same median 
but number of users is more variable on Thursday. (sometimes it is high and sometimes it is very low) 
* Wednesday and Friday have the same
median but number of users is more variable on Friday.

* Monday has the lowest median and it is less variable than other
    workdays (the least number of Active users is on Monday)
    and there is notable difference between the number of users on Monday and any other day  
    
* Saturday has the highest median but the number of users is the least variable (number of users on Sat differ greatly (higher) than Tuesday and Wednesday) 

* Sunday has lower median than Saturday but the number of users is more variable 

in conclusion: the number of users start increasing from Tuesday to Friday then on Saturday it the is highest and mostly stable (not variable) also on Sunday it start decreasing until it reaches the lowest number of users by Monday 

** as the data is normally distributed we can test the following hypothesis 

H: some days have greater number of users than others 
(number of users is associated with weekdays)

H0: There is no  association between number of Active users and weekdays.

---------------

due to the small sample size we need to combine some days in groups

labeling Tuesday-Friday as Tue_Fri, Saturday& Sunday as Sat_Sun 
AND Monday will be on it`s own 

```{r}
weekdays_df_2 <- weekdays_df
weekdays_df_2$day_type <- "W"
weekdays_df_2

Tue_Fri <- c( 
    "Tuesday", "Wednesday", "Thursday", "Friday")
Sat_Sun <- c("Saturday","Sunday")


for (i in 1:31){
  #print(i)
  if (weekdays_df_2$weekday[i] %in% Tue_Fri){
    print(i)
    weekdays_df_2[i,]["day_type"] <- "Tue_Fri"
  }
  if (weekdays_df_2$weekday[i] %in% Sat_Sun){
    weekdays_df_2[i,]["day_type"] <- "Sat_Sun"
  }  
  if (weekdays_df_2$weekday[i] == "Monday"){
    weekdays_df_2[i,]["day_type"] <- "Monday"
  }  
  i = i + 1
}
```



```{r}
weekdays_df_2
```

```{r}
#png("Distribution of number of users.png",
#    width = 800, height = 500)
ggplot(data=weekdays_df_2)+ 
  geom_boxplot(aes(x=day_type,y = user_count),
               outlier.colour="red",
               outlier.shape=8,
               outlier.size=4) +
  xlab("Days")+
  ylab("Number of Active Users")+
  ggtitle("Distribution of number of users")
  #ylim(7,18)
#dev.off()
```
Sat-Sun group has the highest median (hence greater number of users)
Tue-Fri group has lower median than Sat-Sun Group however it is more variable and has an outlier 

Monday has the lowest median (there is huge difference between number of users on Monday and any other day(the lowest)) 

In conclusion: more users (consistent) use the device on weekends (Sat-Sun) than workdays (Tue-Fri). the least number of users is on Monday (consistent)

-----------

We should apply Mann-Whitney U test To test the hypothesis 

###Step 1 
Data preparation 
```{r}
weekdays_test_df <- weekdays_df_2 %>% 
  select(user_count,day_type)
```


```{r}
weekdays_test_df <- weekdays_test_df %>% 
  subset(select= -activity_date)
```


```{r}
weekdays_test_df
```
```{r}
weekdays_test_df %>% group_by(day_type) %>%
  summarise(sum=n())
```

### Step 2
select equal samples 

Sample size = 4 

and for Tue_Fri group only 
sample size = 8 and sample size =4
```{r}
Sat_Sun_subset <- weekdays_test_df %>% 
  filter(day_type == "Sat_Sun")

Tue_Fri_subset <- weekdays_test_df %>% 
  filter(day_type == "Tue_Fri")

Monday_subset <- weekdays_test_df %>% 
  filter(day_type == "Monday")
```

```{r}
Sat_Sun_subset
Tue_Fri_subset  
Monday_subset
```

```{r}
Sat_Sun_sample <- sample_n(Sat_Sun_subset,4)
```

```{r}
Tue_Fri_sample <- sample_n(Tue_Fri_subset,4)
Tue_Fri_sample_8 <- sample_n(Tue_Fri_subset,8)
```

```{r}
Sat_Sun_sample
Tue_Fri_sample
Monday_subset 
Tue_Fri_sample_8
Sat_Sun_subset
```

We have 3 groups (Tue-Fri, Sat-Sun, Monday) so there will be multiple tests for each pair.

**Constructing pairs** 

adding (Sat-Sun) & (Tue-Fri) in one df Sat_Tue_test_df (sample size=8)
adding (Sat-Sun) & (Monday) in one df Sat_Monday_test_df (sample size=4)
adding (Tue-Fri) & (Monday) in one df Tue_Monday_test_df (sample size=4)

```{r}

Sat_Tue_test_df <-rbind(Sat_Sun_subset, Tue_Fri_sample_8)

Tue_Monday_test_df <-rbind(Tue_Fri_sample, Monday_subset) 

```

```{r}
Sat_Monday_test_df <-rbind(Sat_Sun_sample, Monday_subset)
```

```{r}
Sat_Tue_test_df
Sat_Monday_test_df
Tue_Monday_test_df
```
Saving the dataframes as csv to be used in the analysis 

```{r}
Sat_Tue_test_df
Sat_Monday_test_df
Tue_Monday_test_df

#write.csv(Sat_Tue_test_df, "./Data/cleaned_data/Sat_Tue_test_df_28_05_2022_v01.csv",row.names = FALSE)

#write.csv(Sat_Monday_test_df, "./Data/cleaned_data/Sat_Monday_test_df_28_05_2022_v01.csv",row.names = FALSE)

#write.csv(Tue_Monday_test_df, "./Data/cleaned_data/Tue_Monday_test_df_28_05_2022_v01.csv",row.names = FALSE) 

```
read dataframes 

```{r}
Sat_Tue_test_df_csv <- read.csv("./Data/cleaned_data/Sat_Tue_test_df_28_05_2022_v01.csv")
Sat_Monday_test_df_csv <- read.csv("./Data/cleaned_data/Sat_Monday_test_df_28_05_2022_v01.csv")
Tue_Monday_test_df_csv <- read.csv("./Data/cleaned_data/Tue_Monday_test_df_28_05_2022_v01.csv")

```

changing day_type to a factor and labeling the groups 
"Sat_Sun" = 1
"Tue_Fri" = 2
"Monday" = 3
```{r}
#attach(Sat_Tue_test_df_csv)
Sat_Tue_test_df_csv$day_type <- factor(Sat_Tue_test_df_csv$day_type, c("Sat_Sun","Tue_Fri"), labels = c(1,2))
```

```{r}
Sat_Monday_test_df_csv$day_type <- factor(Sat_Monday_test_df_csv$day_type, c("Sat_Sun","Monday"), labels = c(1,3))

```

```{r}
Tue_Monday_test_df_csv$day_type <- factor(Tue_Monday_test_df_csv$day_type, c("Tue_Fri","Monday"), labels = c(2,3))
```

```{r}
str(Sat_Tue_test_df_csv) 
str(Sat_Monday_test_df_csv) 
str(Tue_Monday_test_df_csv) 

```
```{r}
Sat_Tue_test_df_csv
Sat_Monday_test_df_csv
Tue_Monday_test_df_csv
```

 *Pair 1* (Sat_sun & Tue_Fri)

H: There is difference between number of users on Saturday & Sunday compared to the number of users on Tuesday to Friday

H0: There is no difference between number of users on Saturday & Sunday compared to the number of users on Tuesday to Friday

showing summary statistics 

```{r}
Sat_Tue_test_df_csv %>% group_by(day_type)%>% summarise(median_data= median(user_count), iqr=IQR(user_count))
```
group 2 (Tue_Fri) median and iqr for this sample is less than group 1 Sat_Sun 
```{r}
Sat_Tue_test_df_csv
```
```{r}
x_Sat_Sun <- filter(Sat_Tue_test_df_csv,day_type == "1")
x_Tue_Fri <- filter(Sat_Tue_test_df_csv,day_type == "2")

```


```{r}

par(mfrow = c(1,2))
hist(x_Sat_Sun$user_count , main = "Saturday to Sunday", xlab = "Active Users")
hist(x_Tue_Fri$user_count, main = "Tuesday to Friday", xlab = "Active Users")
```
both histograms are positively skewed  
so it is better to use medians to summaries the differences between number of users on (Sat_sun / Tue_Fri) 
if the histograms looks different we should use the mean 

carrying out Mann-Whitney U test 
```{r}
wilcox.test(Sat_Tue_test_df_csv$user_count~Sat_Tue_test_df_csv$day_type)
```
Accepting NULL Hypothesis (H0) for Pair 1 (Sat_sun&Tue_Fri)

H0: There is no difference between number of users on Saturday & Sunday compared to the number of users on Tuesday to Friday

p-value = 0.05299 
Reporting Mann-Whitney U test 

A Mann-Whitney U test showed that there is no significant difference (W = 50.5, p-value = 0.05299) between number of users on Sat_Sun  compared to the number of users on Tue_Fri  (there is no huge difference in the medians of the two groups) 

the median number of users for Sat_Sun group was 14 and Tue_Fri group was 12.5 


*Pair 2* (Sat_sun & Monday)

H: There is difference between number of users on Saturday & Sunday compared to the number of users on Monday 

H0: There is no difference between number of users on Saturday & Sunday compared to the number of users on Monday

showing summary statistics 

```{r}
Sat_Monday_test_df_csv %>% group_by(day_type)%>% summarise(median_data= median(user_count), iqr=IQR(user_count))
```
group 3 (Monday) median and iqr for this sample is less than group 1 Sat_Sun 
```{r}
Sat_Monday_test_df_csv
```

```{r}
x_Sat_Sun_2 <- filter(Sat_Monday_test_df_csv,day_type == "1")
x_Monday <- filter(Sat_Monday_test_df_csv,day_type == "3")

```

```{r}
x_Sat_Sun_2
```


```{r}

par(mfrow = c(1,2))
hist(x_Sat_Sun_2$user_count, main = "Saturday to Sunday", xlab = "Active Users")
hist(x_Monday$user_count, main = "Monday", xlab = "Active Users")
```
both histograms are positively skewed  
so it is better to use medians to summaries the differences between number of users on (Sat_sun / Monday) 

if the histograms looks different we should use the mean 

carrying out Mann-Whitney U test 
```{r}
wilcox.test(Sat_Monday_test_df_csv$user_count~Sat_Monday_test_df_csv$day_type)
```
Accepting NULL Hypothesis (H0) for Pair 2 (Sat_sun&Monday)

H0: There is no difference between number of users on Saturday & Sunday compared to the number of users on Monday

p-value =  0.1081 
Reporting Mann-Whitney U test 

A Mann-Whitney U test showed that there is no significant difference (W = 14, p-value = 0.1081) between number of users on Sat_Sun  compared to the number of users on Monday  (there is no huge difference in the medians of the two groups) 

*Pair 3* (Tue_Fri & Monday)

H: There is difference between number of users on Tuesday to Friday compared to the number of users on Monday 

H0: There is no difference between number of users on Tuesday to Friday compared to the number of users on Monday

showing summary statistics 

```{r}
Tue_Monday_test_df_csv %>% group_by(day_type)%>% summarise(median_data= median(user_count), iqr=IQR(user_count))
```
group 3 (Monday) median and iqr for this sample is less than group 2 Tue_Fri 
```{r}
Sat_Monday_test_df_csv
```

```{r}
x_Sat_Sun_3 <- filter(Sat_Monday_test_df_csv,day_type == "1")
x_Monday <- filter(Sat_Monday_test_df_csv,day_type == "3")

```

```{r}
x_Sat_Sun_3
```

```{r}
x_Sat_Sun_2
```

```{r}

par(mfrow = c(1,2))
hist(x_Sat_Sun_2$user_count, main = "Saturday to Sunday", xlab = "Active Users")
hist(x_Monday$user_count, main = "Monday", xlab = "Active Users")
```
both histograms are positively skewed  
so it is better to use medians to summaries the differences between number of users on (Tue_Fri / Monday) 

if the histograms looks different we should use the mean 

carrying out Mann-Whitney U test 
```{r}
wilcox.test(Tue_Monday_test_df_csv$user_count~Tue_Monday_test_df_csv$day_type)
```
Accepting NULL Hypothesis (H0) for Pair 3 (Tue_Fri&Monday)

H0: There is no difference between number of users on Tuesday to Friday compared to the number of users on Monday

p-value = 1
Reporting Mann-Whitney U test 

A Mann-Whitney U test showed that there is no significant difference (W = 8.5, p-value = 1) between number of users on Tue_Fri  compared to the number of users on Monday  (there is no huge difference in the medians of the two groups) 

*in conclusion:* 
We accept the NULL Hypothesis 
H0: There is no  association between number of Active users and weekdays. 

NOTE: Due to the small sample size 24 (N<30)the results might not be much accurate. We need more data to apply this result on the population. 
